\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2020-02-05 gwnnDraft02.tex
new max plot
made changes suggested by Daniel

The exotic sectors have essentially massless degrees of freedom pre-phase transition and pions with negligible masses post-phase transition. Other composite particles, such as baryons, do gain a mass of the order of $\Lambda_{ex}$; this is, however, still much smaller than the order of $\rho_R$ leading to small $\alpha_{\infty}$ according to Eq.~(\ref{eqn:critPTstrength}).\footnote {It should also be noted that although free quarks cease to exist post phase transition in these exotic sectors, their masses are so light that they do not contribute relevant amounts to $\alpha_{\infty}$.} Heavy standard sectors that undergo SFOPT for QCD feature no baryons due to all quarks being above their respective QCD scales. They do, however, feature glueballs that obtain a mass of the order of the SFOPT and, just as in the case for exotic sectors above, feature small $\alpha_{\infty}$. Ultimately, calculating the critical phase transition strength for each sector using Eq.~(\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. 
%Explicitly, all exotic sectors (and heavy standard sectors) have strong first order phase transitions that lead to runaway bubble walls ($\alpha >\alpha_{\infty}$). 
Each component of the spectral energy density in Eq.~(\ref{eqn:SpectralEng}) is proportional to a power of their respective efficiency factors $\kappa$. The relative strength of the efficiency factors is dependent on the ratio $\frac{\alpha_{\infty}}{\alpha}$ --- since $\alpha_{\infty}$ is small, this would typically lead to the a scenario where the efficiency factors for the sound wave and MHD contributions are small and the GWs produced are dominantly from bubble collisions, $h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi}$: this is what we assume for the remainder of this section. It should be noted, however, that $\alpha$ can be small \cite{Bai:2018dxf,Helmboldt:2019pan}; this leads to the non-bubble collision contributions being important and changes the GW profile. This case and the gravitational waves it produces are examined in the appendix.

@@

Now that the framework has been laid out for the creation of GW from a single SFOPT, we generalize to multiple sectors going under independent, coherent SFOPT. In the models presented in this paper, we consider a subset of $N$ hidden sectors that undergo a phase transition at a SM temperature of $T_{\gamma}^{i}$. As the GWs propagate in free space, the energy density and frequency spectrum, at the time of production $\Omega_{\textrm{GW}}^{*}(f)$, will redshift to today's value $\Omega_{\textrm{GW}}^{0}(f) = \mathcal{A}\, \Omega_{\textrm{GW}}^{*}((a_{0}/a) f) $. The redshifting factor, $\mathcal{A}$, accounts for the redshifting of both $\rho_{\textrm{GW}} $ and $\rho_{c}$~\cite{Breitbach:2018ddu,PhysRevD.49.2837},
\begin{equation}\label{eqn::redshifting}
\mathcal{A} \equiv \bigg( \frac{a}{a_{0}} \bigg)^4 \bigg( \frac{H}{H_{0}}\bigg)^2
\end{equation}
where $a$ ($a_{0}$)  and $H$ ($H_{0}$) are the scale factor and Hubble constant at the time of nucleation (observation), respectively. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW signal that would be measured today is given by the coherent sum,
\begin{equation}
\Omega_{\textrm{GW}} =  \sum_{i}^{N} \mathcal{A}^{i}\, \Omega_{\textrm{GW}}^{i,*}((a_{0}/a)_{i} f) .
\end{equation}
We assume that the parameters of the SFOPT do not differ between sectors: the relativistic degrees of freedom, phase transition rate, and the dark QCD scale, are all similar. This makes the redshifting factor $\mathcal{A}^i$ independent of sector number.
\begin{figure}[h!]
\centering
\includegraphics[scale=.3]{Nnatral.png} 
\caption{ Gravitational wave spectral energy density (solid curves) for standard $N$naturalness using the scalar reheaton model of section~\ref{sec:reheat}. The curve corresponding to the sum of the sectors is approximately equal to the $i=1$ curve. All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$. The colored solid lines use $\beta/H  = 10$ where as the dashed gray line is the total contribution of all sectors for $\beta/H  = 10^4$.  The shaded dashed curves are the power law noise curves~\cite{PhysRevD.88.124032} calculated from expected sensitivity as described in Section~\ref{sec:detection}. The ones on the right are space-based interferometers: Lisa~\citep{Audley:2017drz} (blue), DECIGO~\citep{10.1093/ptep/pty078} (light blue), BBO~\citep{PhysRevD.72.083005} (red). The ones on the left are for the pulsar timing array SKA~\cite{Janssen:2014dka} for exposure time of 5-years (purple), 10-years (orange), and 20-years (green). }
\label{fig::Nnatural}
\end{figure}
Applying this to the standard reheating scenario of $N$natrualness, introduced in~\ref{sec:reheat}, we get GW signals as seen in Fig.~\ref{fig::Nnatural}. Plotted are the individual contributions to the signal from each phase transitioned sector, as well as the coherent sum of all sectors. Future GW interferometers and pulsar timing array sensitivity curves are shown in comparison to the signal. The sensitivity curves are interpreted as the region of possible detection if intersected with the GW signal, and the construction of these curves is detailed in Section~\ref{sec:detection}. Notice that the total signal is dominated by the first sector's contribution. This is caused by the quartic temperature ratio suppression in Eq.~(\ref{eq:alpha}) and the large temperature gaps between adjacent sectors. Such a suppression leads to standard $N$natrualness evading future detector thresholds by a few orders of magnitude in units of energy density.

@@

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=.45\textwidth ]{MaxSW.png}
\hfill
\includegraphics[width=.45\textwidth]{largeSW.png} 
\hfill
\includegraphics[width=.45\textwidth]{MedSW.png} 
\hfill
\includegraphics[width=.45\textwidth]{fiveSW.png} 
\end{minipage}
\hfill
\caption{ Gravitational wave spectral energy density for the various scenarios found in Tab.~(\ref{tab:cases}).  All contributions are assumed to be from a non-runaway phase transition with terminal velocity $v = 0.95$. The colored solid lines use $\beta/H  = 10$ where as the dashed gray line is the total contribution of all sectors for $\beta/H  = 10^4$ . The shaded curves are the same as Fig.~\ref{fig::Nnatural}. In contrast to the runaway case, most scenarios evade the projected sensistivies of future experiments and therefore is the less optimal situation }
\label{fig:Haasw}
\end{figure*}

>>>>

The exotic sectors have essentially massless degrees of freedom pre-phase transition and pions with negligible masses post-phase transition. Other composite particles, such as baryons, do gain a mass of the order of $\Lambda_{ex}$; this is, however, still much smaller than the order of $\rho_R$ leading to small $\alpha_{\infty}$ according to Eq.~(\ref{eqn:critPTstrength}).\footnote {It should also be noted that although free quarks cease to exist post phase transition in these exotic sectors, their masses are so light that they do not contribute relevant amounts to $\alpha_{\infty}$.} Heavy standard sectors that undergo SFOPT for QCD feature no baryons due to all quarks being above their respective QCD scales. They do, however, feature glueballs that obtain a mass of the order of the SFOPT and, just as in the case for exotic sectors above, feature small $\alpha_{\infty}$.

%Explicitly, all exotic sectors (and heavy standard sectors) have strong first order phase transitions that lead to runaway bubble walls ($\alpha >\alpha_{\infty}$). 
Each component of the spectral energy density in Eq.~(\ref{eqn:SpectralEng}) is proportional to a power of their respective efficiency factors $\kappa$. The relative strength of the efficiency factors is dependent on the ratio $\frac{\alpha_{\infty}}{\alpha}$ --- since $\alpha_{\infty}$ is small, this would typically lead to the a scenario with runaway bublle walls where efficiency factors for the sound wave and MHD contributions are small and the GWs produced are dominantly from bubble collisions, $h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi}$: this is what we assume for the remainder of this section. It should be noted, however, that $\alpha$ can also be small \cite{Bai:2018dxf,Helmboldt:2019pan}; this leads to bubbles that are non-runaway (but $v_w \sim 1$ still \cite{Breitbach:2018ddu,Bodeker2017}) such that non-bubble collision contributions being important --- ultimately leading to significant changes to the GW profile. This case and the gravitational waves it produces are examined in the appendix.

@@

Now that the framework has been laid out for the creation of GW from a single SFOPT, we generalize to multiple sectors going under independent, coherent SFOPT. In the models presented in this paper, we consider a subset of $N$ hidden sectors that undergo a phase transition at a SM temperature of $T_{\gamma}^{i}$. As the GWs propagate in free space, the energy density and frequency spectrum, at the time of production $\Omega_{\textrm{GW}}^{*}(f)$, will redshift to today's value $\Omega_{\textrm{GW}}^{0}(f) = \mathcal{A}\, \Omega_{\textrm{GW}}^{*}((a_{0}/a) f) $. The redshifting factor, $\mathcal{A}$, accounts for the redshifting of both $\rho_{\textrm{GW}} $ and $\rho_{c}$~\cite{Breitbach:2018ddu,PhysRevD.49.2837},
\begin{equation}\label{eqn::redshifting}
\mathcal{A} \equiv \bigg( \frac{a}{a_{0}} \bigg)^4 \bigg( \frac{H}{H_{0}}\bigg)^2
\end{equation}
where $a$ ($a_{0}$)  and $H$ ($H_{0}$) are the scale factor and Hubble constant at the time of nucleation (observation), respectively. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW signal that would be measured today is given by the coherent sum,
\begin{equation}
\Omega_{\textrm{GW}} =  \sum_{i}^{N} \mathcal{A}^{i}\, \Omega_{\textrm{GW}}^{i,*}((a_{0}/a)_{i} f) .
\end{equation}
We assume that the parameters of the SFOPT do not differ between sectors: the relativistic degrees of freedom, phase transition rate, and the dark QCD scale, are all similar. This makes the redshifting factor $\mathcal{A}^i$ independent of sector number.
\begin{figure}[h!]
\centering
\includegraphics[scale=.3]{Nnatral.png} 
\caption{ Gravitational wave spectral energy density (solid curves) for standard $N$naturalness using the scalar reheaton model of section~\ref{sec:reheat}. The curve corresponding to the sum of the sectors is approximately equal to the $i=1$ curve. All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$. The colored solid lines use $\beta/H  = 10$ where as the dashed gray line is the total contribution of all sectors for $\beta/H  = 10^4$ (the sum of all sectors is roughly equal to the $i = 1$ curve and sectors beyond the first are below the range of this plot).  The shaded dashed curves are the power law noise curves~\cite{PhysRevD.88.124032} calculated from expected sensitivity as described in Section~\ref{sec:detection}. The ones on the right are space-based interferometers: Lisa~\citep{Audley:2017drz} (blue), DECIGO~\citep{10.1093/ptep/pty078} (light blue), BBO~\citep{PhysRevD.72.083005} (red). The ones on the left are for the pulsar timing array SKA~\cite{Janssen:2014dka} for exposure time of 5-years (purple), 10-years (orange), and 20-years (green). }
\label{fig::Nnatural}
\end{figure}
Applying this to the standard reheating scenario of $N$natrualness, introduced in~\ref{sec:reheat}, we get GW signals as seen in Fig.~\ref{fig::Nnatural}. Plotted are the individual contributions to the signal from each phase transitioned sector, as well as the coherent sum of all sectors. Future GW interferometers and pulsar timing array sensitivity curves are shown in comparison to the signal. The sensitivity curves are interpreted as the region of possible detection if intersected with the GW signal, and the construction of these curves is detailed in Section~\ref{sec:detection}. Notice that the total signal is dominated by the first sector's contribution. This is caused by the quartic temperature ratio suppression in Eq.~(\ref{eq:alpha}) and the large temperature gaps between adjacent sectors. Such a suppression leads to standard $N$natrualness evading future detector thresholds by a few orders of magnitude in units of energy density.

@@

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=.45\textwidth ]{MaxSW.png}
\hfill
\includegraphics[width=.45\textwidth]{largeSW.png} 
\hfill
\includegraphics[width=.45\textwidth]{MedSW.png} 
\hfill
\includegraphics[width=.45\textwidth]{fiveSW.png} 
\end{minipage}
\hfill
\caption{ Gravitational wave spectral energy density for the various scenarios found in Tab.~(\ref{tab:cases}).  All contributions are assumed to be from a non-runaway phase transition with terminal velocity $v = 0.95$. The colored solid lines use $\beta/H  = 10$ where as the dashed gray line is the total contribution of all sectors for $\beta/H  = 10^4$ . The top left figure shows the individua sound wave and MHD contributions.  The shaded curves are the same as Fig.~\ref{fig::Nnatural}. In contrast to the runaway case, most scenarios evade the projected sensistivies of future experiments and therefore is the less optimal situation }
\label{fig:Haasw}
\end{figure*}
